---
title: "1141-modi"
author: "MSPC"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    toc: yes
    toc_float: false
    self_contained: true
params:
  data_dir: NULL
  metadata: NULL
  config: "None"  # from python. make all the NULLs "None"?
  modi_abbrev: ac # not used?
---

<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
</style>

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(rlang))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(reactable))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(here))
library(dendsort)


library(knitr)
myreactable <- function(x, ...) reactable(x, defaultPageSize = 14, searchable = T, compact = T, resizable=T, sortable = T, filterable = T, ...)

(here())

OUTDIR <- params$outdir %||% params$data_dir %||% "."

util_tools <- new.env()
source( file.path(here("R", "utils.R")), local = util_tools )
dist_no_na <- util_tools$dist_no_na
myzscore <- util_tools$myzscore
io_tools <- new.env()
source( file.path(here("R", "io.R")), local = io_tools )

# more fine grained params with defaults and validation
config <- io_tools$get_config(params$config)

```

```{r, label=funcs, include=F}





read_expr <- function(.x){
  df <- readr::read_tsv(.x,
                        #  n_max=4000,
                        show_col_types = FALSE
                        ) %>%
    mutate(expr_file = .x)
  # mutate(modi_id = basename(id) %>% str_extract(pattern = "(sty|k)_[0-9]+_[0-9]+")) %>%

  df %<>% mutate(sitename = if_else(is.na(sitename2), sitename, sitename2))
  if ("geneid" %in% colnames(df)) df %<>% mutate(geneid = as.character(geneid))

  if ("cst_cat#" %in% colnames(df)) df %<>% mutate(`cst_cat#` = as.character(`cst_cat#`))

  if ("all_possible_positions" %in% colnames(df)) df$all_possible_positions <- NULL
  if ("all_possible_positions_psp" %in% colnames(df)) df$all_possible_positions_psp <- NULL

  if ("ms_lit" %in% colnames(df)) df %<>% mutate(`ms_lit` = replace_na(ms_lit, 0))
  if ("lt_lit" %in% colnames(df)) df %<>% mutate(`lt_lit` = replace_na(lt_lit, 0))

  if (!"fifteenmer" %in% colnames(df)) stop('fifteenmer missing')
  if (!"ENSP" %in% colnames(df)) stop('ENSP missing, this is a protein id ')
  if (!"sitename" %in% colnames(df)) stop('sitename missing')
  if (!"taxon" %in% colnames(df)) df$taxon <- "<NULL>"
  df %<>% mutate(site_id = str_c(sitename, ENSP, fifteenmer, sep='_'))

  return(df)
}

```


# load


params:
```{r, label=load_meta}
(params)

if (!is.null(params$metadata)){
    meta <- read_tsv(params$metadata, show_col_types = FALSE)
    if (!"group" %in% colnames(meta)) meta$group <- meta$name
} else {
  cat("no metadata passed, exiting")
  knitr::knit_exit()
}

```

```{r, label = view_meta}
meta %>% myreactable
```


```{r, label=load_expr}

if (!"expr_file" %in% colnames(meta)){
  cat("not implemented yet, exiting")
  knitr::knit_exit()
}

expr_files <- meta$expr_file %>% unique

datas <- expr_files %>%
    purrr::map( read_expr )  %>%
    purrr::set_names(expr_files)

( names(datas) )

```


```{r, label='combine data'}

data <- bind_rows(datas) # %>% dplyr::filter(!str_detect(protein, "\\|Cont_"))

# data %<>% distinct(sitename, ENSP, parent, fifteenmer, .keep_all = T)
# data %<>% distinct(sitename, parent, fifteenmer, hyperscore_best, .keep_all = T)

dplyr::glimpse(data)

# data %>% head() %>% myreactable()

```


# explore

```{r, label='explore'}
data %>%
  group_by(expr_file) %>%
  summarize(n = n()) %>%
  arrange(-n)
```

```{r, label='explore2'}
datal <- data %>%
  pivot_longer(
    cols = starts_with("TMT"), names_to = "expr_col"
  ) %>%
  left_join(meta)

# datal %<>% dplyr::filter(!is.na(name))
datal %>% dplyr::glimpse()
```

# normalize

```{r, label='normalize', cache=F}
# cacheing this is not working in Rmd
.minval <- datal %>%
  dplyr::filter(value > 0) %>%
  pull(value) %>%
  min()
datal %<>%
  group_by(name, recno, runno, searchno, label, taxon) %>%
  mutate(
    val_mednorm = (value + (.minval / 2)) / median(value, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    log2_val = log2(val_mednorm)
  )  %>%
  group_by(site_id) %>% mutate(zscore = util_tools$myzscore(log2_val)) %>% ungroup()

counts <- datal %>%
  dplyr::filter(!is.na(log2_val)) %>%
  group_by(site_id) %>%
  summarize(n=n()) %>%
  arrange(-n)
maxn <- min(max(counts$n), nrow(metadata))
tokeep <- counts %>% filter(n == maxn)
datal %<>% dplyr::filter(
  site_id %in% tokeep$site_id
)

  # (counts)
  # (maxn)
# datal %>% head() %>% myreactable()
```


```{r, eval=T, label='batch correction'}

do_combat <- function(datal){
  data_wide <- datal %>%
    pivot_wider(id_cols = site_id, names_from = name, values_from = log2_val)
  mat <- data_wide[, meta$name] %>% as.matrix
  rownames(mat) <-  data_wide$site_id
  corrected_data <- sva::ComBat(dat = mat, batch = meta$recno %>% as.character )
}
if (!is.null(params$batch)){
  combat_corrected <- do_combat(datal)
  combat_corrected %<>% as.data.frame %>% rownames_to_column(var="site_id")
  combat_correctedl <- combat_corrected %>% pivot_longer(-site_id, values_to = 'log2_val')
  datal2 <- datal %>% left_join(combat_correctedl)
  datal <- datal2
  }

```

```{r, label='inspect data'}
( datal %>%   nrow() )
( datal %>%  distinct(spectra, .keep_all=T) %>% nrow() )
( datal %>%  distinct(name, spectra, .keep_all=T) %>% nrow() )
```

## distributions

```{r}
datal %>%
  distinct(name, spectra, .keep_all=T) %>%
  ggplot(aes(x = log2_val, col = name)) +
  geom_density() +
  facet_wrap(~group)
```


```{r}
datal %>%
  distinct(name, spectra, .keep_all=T) %>%
  ggplot(aes(x = log2_val, col = name)) +
  geom_density() +
  facet_wrap(~expr_file)
```



```{r}
datal %>%
  distinct(name, spectra, .keep_all=T) %>%
  ggplot(aes(x = log2_val, col = name)) +
  geom_density() +
  facet_wrap(~expr_file+AA)
```




```{r, label='pivot_wider'}
# .id_cols <- c("fifteenmer", "ENSG", "ENSP", "geneid", "taxon", "position", "sitename", "id", "modi_id")
#datal %>% dplyr::summarise(n = dplyr::n(), .by = c(fifteenmer, ENSG, ENSP, geneid, taxon, sitename, id, name))
dataw <- datal %>%
  pivot_wider(
    id_cols = site_id,
    names_from = name,
    #values_fn = function(x) median(x, na.rm = T),
    values_from = log2_val, # sometimes this happens after merging with phosphositeplus table
  )
```
# limma
```{r, label='limma'}
library(limma)

mod <- model.matrix(config$limma$formula %||% as.formula('~0+group'), meta)
rownames(mod) <- meta$name
mat <- dataw[, meta$name]
colnames(mod) <- make.names(colnames(mod))

generate_contrasts <- function(df, group_var) {
  # Get unique levels of the grouping variable
  group_levels <- unique(df[[group_var]])

  # Generate all pairwise combinations of group levels
  contrast_combinations <- combn(group_levels, 2, simplify = FALSE)

  # Create contrasts for each combination
  contrast_strings <- lapply(contrast_combinations, function(combo) {
    paste(paste0(group_var, combo[1]), "-", paste0(group_var, combo[2]))
  }) %>% unlist()
  return(contrast_strings)
}


contrast_strings <- config$limma$contrasts %||%  generate_contrasts(meta, "group")
contrasts <- makeContrasts(contrasts = contrast_strings, levels = mod)

 # ( contrasts <- limma::makeContrasts(contrast_strings, levels = mod) )
```






```{r, label='lmFit', cache=TRUE}
fit <- limma::lmFit(mat, mod)
fit <- contrasts.fit(fit, contrasts)
fitres <- fit %>% limma::eBayes(robust = T, trend = T)
```



```{r, label='topTables'}
toptables <- colnames(contrasts) %>%
  purrr::map(~ {
    limma::topTable(fitres, coef = .x, number = Inf, sort.by = "none")
  }) %>%
  setNames(nm = colnames(contrasts))


toptables2 <- toptables %>% purrr::map(~ {
  assertthat::are_equal(nrow(.x), nrow(dataw))
  bind_cols(
    dataw %>% select(-meta$name),
    .x
  ) %>% arrange(P.Value) %>%
    left_join(data %>% select(
      acc_id,
      lt_lit,
      ms_lit,
      fifteenmer,
      sitename,
      uniprot_id,
      ENSP,
      site_id
    ) %>%
    distinct()
  )

})


toptables2 %>% purrr::iwalk(~ {
  .outdir <- file.path(OUTDIR, "results", "modi")
  if (!fs::dir_exists(.outdir)) fs::dir_create(.outdir)
  .outf <- .y %>% str_replace(" - ", "_minus_")
  .fulloutf <- file.path(.outdir, paste0(make.names(.outf), ".tsv"))

  if (fs::file_exists(.fulloutf) && config$advanced$replace == FALSE) {
    cat(paste0("Skipping ", .fulloutf, " because it exists and replace is FALSE\n"))
    return()
  }
  cat(paste0("Writing ", .fulloutf,'\n'))
  .x %>% write_tsv(.fulloutf)

})
```



## stat overview

```{r, label='p value dist'}
plts <- toptables2 %>%
  purrr::imap(
    ~ {
      plt <- .x %>%
        ggplot(aes(x = P.Value)) +
        geom_histogram(binwidth = .05, ) +
        labs(title = paste0(.y, "\n", "p value dist"))
      print(plt)
    }
  )
```


```{r, label='adj p value dist'}
plts <- toptables2 %>%
  purrr::imap(
    ~ {
      plt <- .x %>%
        ggplot(aes(x = adj.P.Val)) +
        geom_histogram(binwidth = .05) +
        labs(title = paste0(.y, "\n", "adj p value dist"))
      print(plt)
    }
  )
```

```{r, include=F}
# to initialize the js dependency
# may not be necessary
reactable(mtcars)
```


```{r, results='asis', eval=T}
._ <- toptables2 %>%
  purrr::imap(~ {
    print(paste0("## ", .y))
    .x %>%
      arrange(P.Value) %>%
      select( -AveExpr, -B) %>%
      select(sitename, everything()) %>%
      head(40) %>%
      myreactable(
        showSortable = T,
        wrap = F,
        bordered = TRUE,
        columns = list(
          sitename = colDef(sticky = "left", style = list(backgroundColor = "#8af")),
          `P.Value` = colDef(format = colFormat(digits = 4)),
          `adj.P.Val` = colDef(format = colFormat(digits = 4)),
          `logFC` = colDef(format = colFormat(digits = 4)),
          `t` = colDef(format = colFormat(digits = 4))
        )
      ) %>%
      htmltools::tagList() %>%
      print()
  })
```



```{r, label = exit_early, eval=F}
save.image(file.path(params$data_dir, 'env.RData'))
knitr::knit_exit()
```


```{r, fig.width=9, fig.height=12}
condense_groups <- function(.mat){

   # Identify numeric and non-numeric columns
  numeric_cols <- sapply(.mat, is.numeric) & !names(.mat) %in% c("ms_lit", "lt_lit")
  non_numeric_cols <- !numeric_cols
  # Group by all numeric columns and summarize non-numeric columns
  .mat2 <- .mat %>%  select(non_numeric_cols[ non_numeric_cols == TRUE] %>% names, numeric_cols[ numeric_cols == TRUE] %>% names )

  numeric_cols <- sapply(.mat2, is.numeric) & !names(.mat2) %in% c("ms_lit", "lt_lit")
  non_numeric_cols <- !numeric_cols

  condensed_data <- .mat2 %>%
    group_by(across(which(numeric_cols))) %>%
    summarize(across(which(non_numeric_cols), ~paste(sort(unique(.)), collapse="/")), .groups = 'drop')
    #summarize(across(which(non_numeric_cols), ~paste(unique(.), collapse="/")), .groups = 'drop')
  #%>%

  condensed_data$sitename %<>% str_replace_all('/', ' / ')

  return(condensed_data)
}

do_plot <- function(.x, .y){

   n <- config$limma$top_n %||% 60

  .sel <- .x %>%
    arrange(P.Value) %>%
    head(n = as.integer(100*n)) # hack

  .data <- datal %>%
    dplyr::filter(site_id %in% .sel$site_id) %>%
    dplyr::filter(!str_detect(name, "Refmix")) # %>%


  # mutate(position_absolut = as.character(position_absolut))
  # .id_cols <- c(
  #   "fifteenmer", "ENSG", "ENSP", "geneid", "symbol", "taxon", "AA", "position_absolut", "sitename", "id", "lt_lit",
  #   "ms_lit", "site_id"
  # )

  .mat <- .data %>% pivot_wider(id_cols = site_id, names_from = name, values_from = zscore, )
  .mat %<>% left_join(datal %>% distinct(site_id, ms_lit, lt_lit, sitename))
  .mat %<>% mutate(ms_lit = replace_na(ms_lit, 0), lt_lit = replace_na(lt_lit, 0))
  .mat <- condense_groups(.mat) %>% head(n = n) # condensing should maybe happen earlier

  # .mat <- .data %>% pivot_wider(id_cols = .id_cols, names_from = name, values_from = zscore, )
  .cids <- intersect(colnames(.mat), meta$name)
  .meta <- meta %>% dplyr::filter(name %in% .cids)

  .colors <- util_tools$create_named_color_list(.meta, c('group'))
  ca <- ComplexHeatmap::columnAnnotation(
    # treat = .meta$treat,
    group = .meta$group,
    annotation_name_side = "left",
    col = .colors
  )

  #
  hide_zero <- function(x) {
    sapply(x, function(y) ifelse(y == 0, "", as.character(y)))
  }

  row_annotations <- HeatmapAnnotation(
    lt_lit = .mat$lt_lit %>% hide_zero() %>% str_wrap(11, whitespace_only=FALSE) %>% anno_text(show_name = T, gp = gpar(fontsize = 7), location = unit(.5, "npc"), just = "centre"),
    ms_lit = .mat$ms_lit %>% hide_zero() %>% str_wrap(11, whitespace_only=FALSE) %>% anno_text(show_name = T, gp = gpar(fontsize = 7), location = unit(.5, "npc"), just = "center"),
    which = "row",
    gp = gpar(fontsize = 2, size = 2),
    width = unit(.72, "in"),
    annotation_name_side = "top",
    border = T,
    show_annotation_name = T
  )

  .ht_mat <- .mat[, .meta$name] %>% as.matrix()
  ht <- ComplexHeatmap::Heatmap(
    .ht_mat,
    width = unit(ncol(.ht_mat) * .2, 'in'),
    height = unit(nrow(.ht_mat) * .24, 'in'),
    column_split = .meta$group,
    row_labels = .mat$sitename %>% stringr::str_wrap(width = 48),
    row_names_gp = grid::gpar(fontsize = 6),
    column_names_side = "top",
    heatmap_legend_param = list(title = "zscore"),
    column_title = .y,
    right_annotation = row_annotations,
    top_annotation = ca,
    cluster_rows = dendsort(hclust(dist(.ht_mat), method = "ward.D2")),
    # cluster_columns = dendsort(hclust(dist(t(.ht_mat)), method="ward.D2"))
    clustering_method_columns = "ward.D2"
    # clustering_method_rows = "ward.D2",
    # clustering_method_columns = "ward.D2"
  )

  .outdir <- file.path(OUTDIR, "results", "modi")
  .outf <- .y %>% str_replace(" - ", "_minus_")
  .fulloutf <- file.path(.outdir, paste0(make.names(.outf), "_top", as.character(n), ".pdf"))
  draw(ht)

  .height <- 7 + (nrow(.ht_mat) * .2 )
  .width <- 9 + (ncol(.ht_mat) * .2 )

  pdf(.fulloutf, height = .height, width = .width)
  print(ht)
  dev.off()


  # same thing but only the two comparison
  .comp <- .y %>%
    str_extract_all("(?<=group)[a-zA-Z0-9]+") %>%
    unlist()
  # .data2 <- .data %>% dplyr::filter(group %in% .comp)
  # .id_cols <- c("fifteenmer", "ENSG", "ENSP", "geneid", "symbol", "taxon", "AA", "position_absolut", "sitename", "id")
  # .mat2 <- .data2 %>%
  #   #group_by(!!!rlang::syms(.id_cols)) %>%
  #   group_by(site_id) %>%
  #   mutate(zscore = util_tools$myzscore(log2_val)) %>%
  #   ungroup() %>%
  #   # pivot_wider(id_cols = .id_cols, names_from = name, values_from = zscore, )
  #   pivot_wider(id_cols = site_id, names_from = name, values_from = zscore, )
  # .mat2 %<>% left_join(datal %>% distinct(site_id, ms_lit, lt_lit, sitename))
  # .mat2 %<>% mutate(ms_lit = replace_na(ms_lit, 0), lt_lit = replace_na(lt_lit, 0))
  # .mat2 <- .mat2 %>% condense_groups() %>% head(n = n)

  .cids <- .data %>% dplyr::filter(group %in% .comp) %>% pull(name) %>% unique()
  .meta2 <- meta %>% dplyr::filter(name %in% .cids)
  .to_remove <- intersect( setdiff(colnames(.mat), .meta2$name), .meta$name)
  .ids <- setdiff(colnames(.mat), .to_remove)
  .mat2 <- .mat[, .ids]
  .cids <- intersect(colnames(.mat2), meta$name)

  .ht_mat2 <- .mat2[, .meta2$name] %>% as.matrix()

  (all(.mat2$sitename == .mat$sitename))
  # .meta2 <- .meta %>%
  #   dplyr::filter(name %in% colnames(.mat2)) %>%
  #   mutate(name = factor(name, levels = colnames(.mat2), ordered = T)) %>%
  #   arrange(name)


  ca <- ComplexHeatmap::columnAnnotation(
    # treat = .meta2$treat,
    group = .meta2$group,
    annotation_name_side = "left",
    col = .colors
  )

  row_annotations <- HeatmapAnnotation(
    lt_lit = .mat2$lt_lit %>% hide_zero() %>% anno_text(show_name = T, gp = gpar(fontsize = 7), location = unit(.5, "npc"), just = "centre"),
    ms_lit = .mat2$ms_lit %>% hide_zero() %>% anno_text(show_name = T, gp = gpar(fontsize = 7), location = unit(.5, "npc"), just = "center"),
    gp = gpar(fontsize = 2, size = 2),
    width = unit(.72, "in"),
    annotation_name_side = "top",
    border = T,
    show_annotation_name = T,
    which = "row"
  )

  ht <- ComplexHeatmap::Heatmap(
    .ht_mat2,
    column_split = .meta2$group,
    width = unit(ncol(.ht_mat2) * .2, 'in'),
    height = unit(nrow(.ht_mat2) * .24, 'in'),
    row_labels = .mat2$sitename %>% stringr::str_wrap(width = 38),
    row_names_gp = grid::gpar(fontsize = 6),
    column_names_side = "top",
    heatmap_legend_param = list(title = "zscore"),
    column_title = .y,
    top_annotation = ca,
    right_annotation = row_annotations,
    cluster_rows = dendsort(hclust(dist(.ht_mat2), method = "ward.D2")),
    # cluster_columns = dendsort(hclust(dist(t(.ht_mat2)), method="ward.D2"))
    clustering_method_columns = "ward.D2"
  )

  .outdir <- file.path(OUTDIR, "results", "modi")
  .outf <- .y %>% str_replace(" - ", "_minus_")
  .fulloutf <- file.path(.outdir, paste0(make.names(.outf), "_top", as.character(n), "_subsel_", nrow(.ht_mat2), 'x', ncol(.ht_mat2), ".pdf"))
  draw(ht)

  .height <- 7 + (nrow(.ht_mat2) * .2 )
  .width <- 8 + (ncol(.ht_mat2) * .2 )
  pdf(.fulloutf, height = .height, width = .width)
  print(ht)
  dev.off()

}

toptables2 %>% purrr::imap(~ do_plot(.x, .y) )
